---
title: "HPC_Assignment1"
author: "HPC"
date: "7/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Peer-graded Assignment: Course Project 1

##First step: Create the dataset

This segment has the code for reading in the dataset and processing the data.

```{r}
library(readr)
activity <- read_csv("activity.csv")
str(activity)
```

## Second step: Summary statistics for the data

What is the mean and median of total number of steps taken per day?

```{r}
library(dplyr)
activity_daily <- activity %>%
                    group_by(date) %>%
                    summarise(daily_steps = sum(steps))
summary(activity_daily)
mean_daily <- mean(activity_daily$daily_steps, na.rm=TRUE)
median_daily <- median(activity_daily$daily_steps, na.rm=TRUE)
```

The mean of the total number of steps taken per day is `mean_daily`.
The median of the total number of steps taken per day is `median_daily`.

And here is the histogram:

```{r}
library(ggplot2)
ggplot(activity_daily, aes(x=date, y=daily_steps)) + geom_bar(stat='identity')

```

And here is a plot with the daily activity pattern:

```{r}
activity_avgday <- activity %>%
                    group_by(interval) %>%
                    summarise(interval_steps = mean(steps, na.rm = TRUE))
ggplot(activity_avgday, mapping = aes(x = interval, y = interval_steps)) + geom_line()


```


##Third step: Inputing missing values

Here we diagnose the missing values problem and create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
count_NAs <- sum(is.na(activity_daily$daily_steps))
```

There exist `count_NAs` missing values.
To fix the issue, we will input mean values where there is a missing value.
Afterwards, we calculate the mean, median and plot an histogram.


```{r}
library(tidyr)
activity_daily2 <- activity_daily %>%
                    replace_na(list(daily_steps=mean_daily))
```

```{r}
summary(activity_daily2)
mean_daily_2 <- mean(activity_daily2$daily_steps, na.rm=TRUE)
median_daily_2 <- median(activity_daily2$daily_steps, na.rm=TRUE)

ggplot(activity_daily2, aes(x=date, y=daily_steps)) + geom_bar(stat='identity')

```

In sum: the strategy used for removing the NAs did not impact the mean but it impacted the median.
- New mean is `mean_daily_2` same as old mean (`mean_daily`)
- New median is `median_daily_2` different from old median (`median_daily`)


##Fourth step: Activity patterns in weekdays vs weekends

Create the new dataset: 

```{r}
activity2 <- activity %>%
                    replace_na(list(steps=mean_daily)) %>%
                    mutate(weekday = weekdays(date),
                           weekday_dummy = ifelse(weekday == "Saturday" | weekday == "Sunday", "Weekend", "Weekday"))
activity2$weekday_dummy <- factor(activity2$weekday_dummy)

str(activity2)

```

Plot the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
activity_avgweekday <- activity2 %>%
                    filter(weekday_dummy == "Weekday") %>%
                    group_by(interval) %>%
                    summarise(interval_steps = mean(steps, na.rm = TRUE))

ggplot(activity_avgweekday, mapping = aes(x = interval, y = interval_steps)) + 
  geom_line()

activity_avgweekend <- activity2 %>%
                    filter(weekday_dummy == "Weekend") %>%
                    group_by(interval) %>%
                    summarise(interval_steps = mean(steps, na.rm = TRUE))

ggplot(activity_avgweekend, mapping = aes(x = interval, y = interval_steps)) + 
  geom_line()


```